Verifying Demonstrations

### 2018-11-01 paper/DB-anon-cred.tex
Minor fixes in DB-anon-cred

During one round, the prover commits to a random nonce: \ie chooses 
\(\rho\rgets \ZZ_q\) uniformly randomly, computes \(R\gets g^\rho\) and sends 
\(R\) to the verifier.
The verifier generates two challenges \(c_0\rgets \bin^k, c_1\rgets \bin^k\) 
and sends them to the prover.
The prover computes \(s_0\gets \rho - c_0\alpha, s_1\gets \rho - c_1\alpha\).
This is the main difference to the original Schnorr protocol: the verifier 
selects two challenges and the prover \emph{computes} two responses --- but 
still only \emph{one nonce}, \(\rho\).
(This is also different from Brands-Chaum, where the prover and verifier 
jointly construct \emph{one} challenge with \emph{one} response.)

Now the prover randomly chooses
\(s_0^{(0)}, s_1^{(0)}\in \ZZ_q\)
and computes
\(s_0^{(1)} \gets s_0 - s_0^{(0)}, s_1^{(1)} \gets s_1 - s_1^{(0)}\).
\Ie \(s_b, b\in\bin\) can be reconstructed from the two shares \(s_b^{(0)}, 
  s_b^{(1)}\).

We let \(|q| = \ceil{\log_2 q}\) denote the length of \(q\) in bits.
The verifier will request all \(2|q|\) response-bits from one challenge (say 
\(s_b^{(0)}\concat s_b^{(1)}, b\in \bin\)) and only \(0 < l\leq |q|\) from the 
other (\(s_{1-b}^{(0)}\concat s_{1-b}^{(1)}\)).
Then the verifier can compute \(s_b = s_b^{(0)} + s_b^{(1)}\) and finally 
authenticate the prover by checking if \(R = g^{s_b}A^{c_b}\).

The prover must know the responses for both challenges to successfully pass the 
\ac{DB} phase.
The reason for the two challenges but only one random nonce (change from 
Schnorr) is that knowing the responses to both means learning \(\alpha\) (for
\ac{DBTF}).
Bundling the authentication into the distance bounding phase (difference from 
Brands-Chaum) was done for \ac{DBDH},
and both adaptations together aim at \ac{DBMF} resistance. %reducing success of MF according to number of repetitions.
The reason we use secret sharing and not \(s_b\) directly, is to bound the 
\(l\) bits the verifier must learn of \(s_{1-b}\):
Without secret sharing, the verifier learns \(l\) bits of \(s_{1-b}\), but with 
secret sharing the verifier learns \(l\) bits of \emph{one share} (provided 
\(l\) is small enough).
Thus we have information theoretic security although we reveal these \(l\) 
bits.

@@

@@

The \ac{DB} phase also ensures that \iac{DBMF} adversary will fail the \ac{DB} 
phase for at least one round.
The verifier will send \(|q| + l < 2|q|\) challenges.
By replying, the prover reveals all the bits of either \(s_0\) or \(s_1\), but 
none of the other --- only \(l\leq |q|\) bits of \emph{one of two} information 
theoretically independent \(|q|\)-bit shares.
The adversary now has two options.
First, the adversary can buffer \(l\) bits of both \(s_0, s_1\) by requesting 
them from the prover.
However, at this point the adversary must wait for the challenge bit 
\(b_{l+1}\) from the verifier and then relay that challenge to the prover to 
receive the correct \(r_i\) --- this relay will be detected.
The (better) alternative is to guess \(b\) and extract \(s_b\) for this 
challenge from the prover.
Guessing \(b\) yields \(1/2\) probability \emph{per round} that the \ac{DBMF} 
adversary will guess it correctly.
Since we must execute the protocol \(n\) times, this is reduced to \(2^{-n}\).
The adversary can also guess the remaining \(2|q|-l\) bits, which we will 
discuss more below.
The probability of guessing these bits successfully is \(2^{-(2|q|-l)}\).
However, \(l < |q| \implies 2|q|-l > |q|\) and this might thus be more 
expensive than guessing \(s_{1-b}\) directly, which is \(2^{-|q|}\).
Thus the success probability of the \ac{DBMF} adversary is \emph{at most} 
\(2^{-n}+2^{-|q|n}\).

@@

The whole Tamarin code we wrote is provided in Appendix \ref{apdx:tamarin-spec}. It contains a specification of the distance-bounding protocol described previously, a definition of the soundness and correctness properties that ensure the protocol actually produces the expected result, and a definition of the security properties we want to ensure.

Tamarin auto-prover validated the soundness and correctness properties (in negligible computation time) but does not terminate when trying to prove the security properties, that is to say, no trace violating the security properties is found, but the prover cannot conclusively validate that the properties are true. We are currently investigating with Tamarin interactive mode to hopefully validate the security properties as well.
\simon{test by sonja}
%\sonja{\subsection{Formal Verification} 
%  verification with tamarin: there is no BC Schnorr in Tamarin,  [10]
% assumed equivalence with Fiat-Shamir model. Not straightforward to
% implement Schnorr in Tamarin due to a lack of expressiveness for
% arithmetic needed to complement Diffie-Hellman. If not done now, left
% for future work. }

>>>>

During one round, the prover commits to a random nonce: \ie chooses 
\(\rho\rgets \ZZ_q\) uniformly randomly, computes \(R\gets g^\rho\) and sends 
\(R\) to the verifier.
The verifier generates two challenges \(c_0\rgets \bin^k, c_1\rgets \bin^k\) 
and sends them to the prover.
The prover computes \(s_0\gets \rho - c_0\alpha, s_1\gets \rho - c_1\alpha\).
This is the main difference to the original Schnorr protocol: the verifier 
selects two challenges and the prover computes two responses --- but still only 
\emph{one nonce}, \(\rho\).
(This is also different from Brands-Chaum, where the prover and verifier 
jointly construct \emph{one} challenge with \emph{one} response.)

Now the prover randomly chooses
\(s_0^{(0)}, s_1^{(0)}\in \ZZ_q\)
and computes
\(s_0^{(1)} \gets s_0 - s_0^{(0)}, s_1^{(1)} \gets s_1 - s_1^{(0)}\).
\Ie \(s_b\) can be reconstructed by adding the two shares \(s_b^{(0)}, 
  s_b^{(1)}\).

We let \(|q| = \ceil{\log_2 q}\) denote the length of \(q\) in bits.
The verifier will request all \(2|q|\) response-bits from one challenge (say 
\(s_b^{(0)}\concat s_b^{(1)}\)) and only \(0 < l\leq |q|\) from the other 
(\(s_{1-b}^{(0)}\concat s_{1-b}^{(1)}\)).
Then the verifier can compute \(s_b = s_b^{(0)} + s_b^{(1)}\) and finally 
authenticate the prover by checking if \(R = g^{s_b}A^{c_b}\).

@@

The prover must know the responses for both challenges to successfully pass the 
\ac{DB} phase.
The reason for the two challenges but only one random nonce is that knowing the 
responses to both means learning \(\alpha\).
This allows us to prevent \ac{DBTF}.
Bundling the authentication into the distance bounding phase (difference from 
Brands-Chaum) prevents \ac{DBDH},
and both adaptations together aim at \ac{DBMF} resistance. %reducing success of MF according to number of repetitions.
The reason we use secret sharing and not \(s_b\) directly, is to bound the 
\(l\) bits the verifier must learn of \(s_{1-b}\):
Without secret sharing, the verifier learns \(l\) bits of \(s_{1-b}\), but with 
secret sharing the verifier learns \(l\) bits of \emph{one share}.
Thus we still have information theoretic security for \(s_{1-b}\) although we 
reveal the \(l\) bits.

@@

The \ac{DB} phase also ensures that \iac{DBMF} adversary will fail the \ac{DB} 
phase for at least one round.
The verifier will send \(|q| + l < 2|q|\) challenges.
By replying, the prover reveals all the bits of either \(s_0\) or \(s_1\), but 
none of the other --- only \(l\leq |q|\) bits of \emph{one of two} information 
theoretically independent \(|q|\)-bit shares.
The adversary now has two options.
First, the adversary can buffer \(l\) bits of both \(s_0^{(0)}\concat 
  s_0^{(1)}, s_1^{(0)}\concat s_1^{(1)}\) by requesting them from the prover.
However, at this point the adversary must wait for the challenge bit 
\(b_{l+1}\) from the verifier and then relay that challenge to the prover to 
receive the correct \(r_i\) --- this relay will be detected.
The alternative is to guess \(b\) and extract \(s_b\) for this challenge from 
the prover.
Guessing \(b\) yields \(1/2\) probability that the \ac{DBMF} adversary will 
guess it correctly.
Since we must execute the protocol \(n\) times, this is reduced to \(2^{-n}\).
The adversary can also guess the remaining \(2|q|-l\) bits, which we will 
discuss more below.
The probability of guessing these bits successfully is \(2^{-(2|q|-l)}\).
However, \(l < |q| \implies 2|q|-l > |q|\) and this might thus be more 
expensive than guessing \(s_{1-b}\) directly, which is \(2^{-|q|}\).
Thus the success probability of the \ac{DBMF} adversary is \emph{at most} 
\(2^{-n}+2^{-|q|n}\).

@@

The whole Tamarin code we wrote is provided in \cref{apdx:tamarin-spec}. It contains a specification of the distance-bounding protocol described previously, a definition of the soundness and correctness properties that ensure the protocol actually produces the expected result, and a definition of the security properties we want to ensure.

Tamarin's auto-prover validated the soundness and correctness properties (in negligible computation time) but does not terminate when trying to prove the security properties, that is to say, no trace violating the security properties is found, but the prover cannot conclusively validate that the properties are true. We are currently investigating with Tamarin interactive mode to hopefully validate the security properties as well.
\simon{test by sonja}
%\sonja{\subsection{Formal Verification} 
%  verification with tamarin: there is no BC Schnorr in Tamarin,  [10]
% assumed equivalence with Fiat-Shamir model. Not straightforward to
% implement Schnorr in Tamarin due to a lack of expressiveness for
% arithmetic needed to complement Diffie-Hellman. If not done now, left
% for future work. }
